### 2015-12-10 paper/paper.tex
Finished preprocessing section

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{latexsym}
\usepackage{multirow}

\title{Title}
\author{Andrew Mason \and Jonathan Pfeil}

\begin{document}
\maketitle
\tableofcontents
\pagebreak

\section{Introduction}
\section{Background}
\subsection{Semantic Representation}

Abstract Meaning Representation (AMR)\cite{amr_sembank} attempts to unify the
semantic representation of natural language sentences. Unlike syntactic
annotations, semantic annotations often differ for different semantic relations
(named entities, temporal entities, co-referencing, etc). AMR, by contrast,
provides a single, graphical representation of the logical meaning of a
sentence, which can be easily understood by both humans and computers. For
example, given the sentence ``A prosecutor in South Korea seized 80 kilograms
of drugs'', we can construct the following logical form and AMR.\\

\noindent \textbf{Logical form}:
\begin{equation}
\begin{split}
\exists s,d2,p2,c,n: &\\
&\text{instance}(s,want\text{-}01) \wedge \text{instance}(d2, drug) \wedge \\
&\text{instance}(p2, prosecutor) \wedge \text{instance}(c, country) \wedge \\
&\text{instance}(n, name) \wedge \text{arg0}(s,p2) \wedge \text{arg1}(s,d2)
\wedge \\ &\text{arg2}(s,c) \wedge \text{name}(c,n) \wedge \text{op1}(n,
``South") \wedge \\ &\text{op2}(n, ``Korea") \wedge \text{unit}(d2, kilogram)
\wedge \text{quant}(d2, 80)
\end{split}
\end{equation}
\textbf{AMR form}:
\begin{figure}
\includegraphics[width=\linewidth]{amr_example.pdf}
\caption{Example AMR}
\label{fig:amr_example}
\end{figure}

An AMR-annotated sentence is directed, rooted, labeled graph, or as a
conjunction of logical triples. Throughout this paper, we work with the AMRs
exclusively in the graphical form $G=(V,E)$, and will thus refer to the AMR
using just the sets $V,E$.

\subsection{Discourse Planning}

In Reiter and Dale's {\em Building Applied Natural Language Generation Systems}, they outline a 6-module pipeline for an NLG system. The second task in this system is Discourse Planning, which they describe as the process of imposing ordering and structure over the set of messages to be conveyed \cite{applied_nlg}. The order of pieces of information in text is not random; rather authors impose a structure on the messages for reasons such as: making the text easier to understand, highlighting the relationships between messages in the text, or stylistic preferences.When making an argument, authors must order the messages so that readers can see the logical implications from one message to the next; if the ordering of the messages were scrambled, readers may fail to understand the argument, even though the same set of information was communicated.

In addition to just logical implication, adjacent sentences can have many different types of relationships with each other, referred to as {\em discourse relations}. A commonly used set of discourse relations are those proposed by {\em Rhetorical Structure Theory} (RST) \cite{rst}. RST defines 32 discourse relations across 3 categories: ``Presentational Relations'', ``Subject Matter Relations'', and ``Multinuclear Relations''.

In applied systems, discourse planning is usually implemented through either schema-based systems or AI planning systems \cite{applied_nlg}. In schema-based systems, messages are given class labels, and are substituted into the appropriate slots which have been predefined in each schema. AI planning is a more general solution in which planning operators are associated with discourse relations, and then off-the-shelf planners can be used to construct a valid discourse plan \cite{hovy1993automated}\cite{paris1990natural}.

\subsection{Information Ordering}

Information ordering is an integral component of the Discourse Planning phase
of the NLG pipeline, discussed above and in \cite{applied_nlg}. After an
information set has been structured into to pieces of information (``facts" or
sentences), these pieces need to be ordered so that the reader can process and
understand the information being conveyed as easily as possible. This is often
approached as a planning problem, though Reiter and Dale explore using decision
trees and instance-based learning.

\section{Related Work}

Early work by Kukich described a system for ``knowledge-based report generation'' in which manually specified domain rules were used to merge database facts into higher level messages, which were then ordered by a discourse module \cite{kukich1983knowledge}. The ordered output could later be passed to generation system to build a report. Work by Paris in 1990 and Hovy in 1993 introduced the idea of using AI Planners to generate discourse plans. They focused on finding explicit discourse relations between messages and using planners to generate an ordering which was allowed under the constraints of the discourse relations identified \cite{paris1990natural}\cite{hovy1993automated}. More recently, Duboue and Mckeown applied Machine Learning (ML) approaches to discourse planning. In \cite{duboue2001empirically}, they present an algorithm to learn ordering constraints among facts. In \cite{duboue2002content}, they build upon this work, using evolutionary algorithms to learn the tree representation of a planner. Although these systems seek to generate valid discourse plans, they do not use a scoring function to choose the `best' among the many valid plans that are available. Dimitromanolaki and Androutsopoulos used ML techniques to learn such a scoring function \cite{learning_to_order_facts}. They train a pipeline of classifiers that repeatedly choose the next fact to be added to a paragraph for a specific domain.

Our work differs from this prior research in three major ways: (1) It is strictly at a semantic level. Our entire pipeline is built upon AMR semantics, rather than natural language `facts', meaning the internal entities and their relations can be reasoned about. (2) Higher-level messages are constructed as part of the discourse planning process. While other systems work at the individual fact level, our system seeks to group facts into higher level sentential messages that are ordered by our discourse module. The order of sentential units should be dependent on the information they contain, therefore the merging procedure must occur before or during the discourse planning. (3) Our system does not explicitly model the discourse relations between units.

\section{Data}

\section{Preprocessing}
Before we can feed our dataset through our featurizer and training/test pipeline, we have to apply various transformations to it: the first two are to ensure uniform sentence semantics and the third is to build our paragraph semantics.

\subsection{ARG-of Reversal}
The typical annotation for a sentence is the one shown in figure \ref{fig:amr_example}. However, due to the flattened textual representation of AMR, it is sometimes easier for annotators to add reversed edges. For example, in figure \ref{fig:amr_example}, instead of adding the edge: (`s', `p2', `ARG0'), they could have added (`p2', `s', `ARG0-of'), resulting in figure \ref{fig:amr_example_argof}. To fix this, we just reverse any such edge and change the label from `ARGN-of' to `ARGN'.

\begin{figure}
\includegraphics[width=\linewidth]{amr_example_argof.pdf}
\caption{AMR with Reversed Edge}
\label{fig:amr_example_argof}
\end{figure}

\subsection{Translation of `and' Semantics}
The next issue that we had to address in the sentence semantics was the way `and' is handled in AMR. For ease of annotation when `and' occurs in a sentence, an `and' instance is created with everything being `anded' together as its children. Then this `and' instance is used as the argument wherever one of the instances children would have gone before. For example, if instead our example sentence were ``A prosecutor in South Korea seized 80 kilograms of drugs and dozens of weapons.'', we would have the AMR graph shown in figure \ref{fig:amr_example_and}.

\begin{figure}
\includegraphics[width=\linewidth]{amr_example_and.pdf}
\caption{AMR with `and' Instance}
\label{fig:amr_example_and}
\end{figure}

While this is representation is not necessarily {\em wrong}, the choice to `and' these two entities together was syntactic, not semantic. A strictly semantic representation of this sentence should contain the information: there is a prosecutor, there are drugs, there are weapons, the prosecutor seized the drugs, and the prosecutor seized the weapons. In a generation process, this could be realized as one sentence (``A prosecutor in South Korea seized 80 kilograms of drugs and dozens of weapons.'') or two sentences (``A prosecutor in South Korea seized 80 kilograms of drugs. The prosecutor also seized dozens of weapons.''). Since we ultimately want to look at how information in paragraphs is grouped into sentences and then ordered, we need to translate these `and' semantics back to a strictly semantic representation; otherwise, we have already encoded some of our information grouping.

To recover the actual semantics from the `and' semantics, we must make N copies of each `and' instance, where N is the number of children of that `and' instance. Then, for each child, we must delete the copied `and' instance and replace it with the child. This algorithm is written formally in algorithm \ref{alg:translate_and_semantics}.

When we apply this algorithm to our semantic graph in figure \ref{fig:amr_example_and}, a new `seize-01' instance is made with the `drugs' instance as ARG1 and the `weapons' instance as ARG1 respectively. This resulting graph is shown in figure \ref{fig:amr_example_and_removed}.

\begin{algorithm}
\caption{Translate `and' Semantics}
\label{alg:translate_and_semantics}
\begin{algorithmic}[1]
%require: amr graph
\REQUIRE AMR Graph $G = (V, E)$

% While the graph still has and instances
\WHILE{$|G.and\_instances| > 0$}
    % and_instance = first and instance in topological ordering of G.and_instances
    \STATE $a \gets$ first and\_instance in topological ordering of $G$.and\_instances

    % S = subgraph formed by performing a breadth first search following parent edges from and_instance upwards to roots
    \STATE $S \gets$ subgraph formed by performing a breadth first search following parent edges from $a$ upwards to roots

    % for each child of the and instance
    \FOR{$c \in$ children($a$)}
        % copy S
        \STATE $S' \gets copy(S)$

        % replace and_instance in S with child
        \STATE replace $a$ with $c$ in $S'$

        % add copy to graph
        \STATE $G \gets G \cup S'$
    \ENDFOR
    % delete and_instance and S from graph
    \STATE remove $a$ and $S$ from $G$
\ENDWHILE
% return G
\RETURN $G$
\end{algorithmic}
\end{algorithm}

\begin{figure}
\includegraphics[width=\linewidth]{amr_example_and_removed.pdf}
\caption{AMR with Translated `and' Instance}
\label{fig:amr_example_and_removed}
\end{figure}

\subsection{Graph Merging}
After the semantic graphs have been preprocessed for each sentence, they have to be merged together to form paragraph semantic graphs. The only complication is deciding if two entities in separate sentences are the same. For example if we are merging the sentences ``The red dog chased the cat. Afterwards, the dog slept.'', the two dogs should be recognized as the same entity. However, if the sentences were ``The red dog chased the cat. The black dog slept.'', the two dogs should be recognized as separate.

We define two nodes $u, v$ as {\b equivalent} if they are two `concept' nodes with the same label, or they have no conflicting attributes and no conflicting children. Formally:

\[
    equiv(u,v) = \left\{\begin{array}{lr}
    u.\text{label} = v.\text{label}, & u,v \text{ `concepts'} \\
    \forall k \text{ attribute of } u \text{ and } v, v.\text{attr}[k] = v.\text{attr}[k] \text{ } \land, & \text{else} \\
    \forall e,f \text{ childedge of } u,v \text{ with same label, } equiv(e.\text{child}, f.\text{child}) & \\
    \end{array}\right\}
\]

To merge two sentences, we use our node equivalence function in addition to algorithm \ref{alg:merge_graphs}. Since this algorithm returns a new, merged graph, we can continually apply it to a list of sentences as a reduction function, resulting in a semantic paragraph graph. Figure \ref{fig:amr_example_3_merged} shows the resultant semantic paragraph graph of the following piece of text: ``A prosecutor in South Korea seized 80 kilograms of drugs. The prosecutor disposed of the drugs. South Korean media stated the drugs originated from North Korea.'' We will continue to use this paragraph as an example throughout the paper.

The only issue of concern during graph merging is what to do when a node in one graph has multiple equivalent nodes in the second graph. This can be caused by annotation errors (i.e. a `name' attribute is annotated ``John F. Kennedy'' in one sentence and ``JFK'' in another) or underspecification (i.e. there are two dogs in the world and the next sentence simply refers to ``the dog''). When this happens, there is no clear resolution, so we drop the errant paragraph from our dataset.

\begin{algorithm}
\caption{Merge Two Sentence Semantic Graphs}
\label{alg:merge_graphs}
\begin{algorithmic}[1]
\REQUIRE Two AMR Graphs $G, H$

\FOR{$u \in H.\text{nodes}$}
    \STATE $equiv\_nodes \gets \{v \mid v \in G.\text{nodes} \land equiv(u,v)\}$
    \IF{$|equiv\_nodes| = 0$}
        \STATE $G.\text{nodes} \gets G.\text{nodes} \cup \{\text{safe\_rename}(G, u)\}$
    \ELSIF{$|equiv\_nodes| = 1$}
        \STATE $equiv\_nodes[0].\text{attr} \gets equiv\_nodes[0].\text{attr} \cup u.\text{attr}$
    \ELSE
        \STATE Raise Ambiguity Error
    \ENDIF
\ENDFOR

\STATE $G.\text{edges} \gets G.\text{edges} \cup H.\text{edges}$

\RETURN $G$
\end{algorithmic}
\end{algorithm}

\begin{figure}
\includegraphics[width=\linewidth]{amr_example_3_merged.pdf}
\caption{Merged AMR Graph of Three Sentences}
\label{fig:amr_example_3_merged}
\end{figure}

\section{System Architecture}
\subsection{Subgraph Selection}
\subsection{Subgraph Ordering}

\section{Evaluation of Scoring Modules}
\subsection{Subgraph Selection}
\subsection{Subgraph Ordering}

\section{Experiments}

\section{Conclusion and Future Work}

\pagebreak
\bibliographystyle{acm}
\bibliography{paper}
\end{document}